---
title: "cluster_outcomes"
output: html_document
author: Simon Aberger
date: "2025-10-23"
---


## Comparison of outcomes and covariates between Clusters

# Import and list relevant libraries
```{r echo=FALSE}
library(survival)
library(survminer)
library(rstatix)
library(tidyverse)
library(cmprsk)
library(MASS)
library(viridis)
```

# Load results
```{r}
load("cluster_results.RData")
```

## Step 1: Kaplan-Meier Survival Analysis (cesnored for KT) between PAM clusters
```{r}
# 1. Get events
oneyear_analysisdata <- oneyear_analysisdata %>%
  mutate(
    time_to_event = pmin(tto_death1, tto_kt1, 365, na.rm = TRUE),
    event = if_else(death_v1 == 1, 1, 0)
  )

# 2. Relabel cluster levels (for legend clarity)
oneyear_analysisdata$cluster_pamGOWER <- factor(
  oneyear_analysisdata$cluster_pamGOWER,
  labels = paste("Cluster", 1:length(unique(oneyear_analysisdata$cluster_pamGOWER)))
)

# and create survival object
surv_objPAM <- Surv(
  time = oneyear_analysisdata$time_to_event,
  event = oneyear_analysisdata$event
)

# 4. Fit KM model
km_fit_pam <- survfit(surv_objPAM ~ cluster_pamGOWER, data = oneyear_analysisdata)

# 5. Plot KM curve
km_plot_pam <- ggsurvplot(
  fit = km_fit_pam,
  data = oneyear_analysisdata,
  palette = c("#1b9e77", "#d95f02", "red", "#7570b3"),
  legend.title = "Cluster",
  legend.labs = levels(oneyear_analysisdata$cluster_pamGOWER),
  risk.table = TRUE,
  xlim = c(0, 365),                         # Limit to 365 days
  break.time.by = 90,                       # Tick marks at 0, 90, 180, 270, 360
  censor = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  xlab = "Time (days)",
  ylab = "Survival Probability",
  title = "Cumulative Survival by Cluster (Censored for KT)",
  risk.table.height = 0.35,
  risk.table.fontsize = 3.5,
  ggtheme = theme_minimal(base_size = 14) +
    theme(
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text = element_text(color = "black"),
      axis.title = element_text(color = "black")
    )
)

km_plot_pam$plot <- km_plot_pam$plot +
  scale_x_continuous(breaks = c(0, 90, 180, 270, 365)) # Adjust x-axis ticks

print(km_plot_pam) # Show plot
```


## Step 2: CIF for secondary outcomes with competing events:
```{r}
# 1. All-cause death
oneyear_analysisdata <- oneyear_analysisdata %>%
  mutate(
    fstatus = case_when(
      death_v1 == 1 ~ 1,             # any cause of death
      transplant_v1 == 1 ~ 0,        # censored at transplant
      TRUE ~ 0
    ),
    ftime = pmin(tto_death1, tto_kt1, 365, na.rm = TRUE)
  )

cuminc_death <- cuminc(ftime   = oneyear_analysisdata$ftime, fstatus = oneyear_analysisdata$fstatus, group   = oneyear_analysisdata$cluster_pamGOWER)
print(cuminc_death)
```

# This process is repeated for `Cause-specific death`, `MACE`, `Interventions for ASCVD`, `Hospitalisation for Heart failure`, `Hospitalisation for infections` and `Fractures`



## Step 3: Incremental prediction model to test impact of cluster membership on the primary endpoint (death)
```{r}
# 1. Prepare Fine-Gray time and event variables
oneyear_analysisdata <- oneyear_analysisdata %>%
  mutate(
    status_fg = case_when(
      death_v1 == 1 ~ 1,  # Death = event
      !is.na(tto_kt1) & tto_kt1 <= 365 ~ 2,  # KT = competing risk
      TRUE ~ 0  # Censored
    ),
    time_fg = pmin(tto_death1, tto_kt1, 365, na.rm = TRUE)
  )

# 3. Add dummy for Cluster 3 (no imputation needed)
finegray_data <- oneyear_analysisdata %>%
  mutate(
    cluster3_vs_rest = {
      cl <- as.character(cluster_pamGOWER)
      as.integer(grepl("(^3$)|3$", cl) | grepl("^Cluster\\s*3$", cl))
    }
  )

# 4. Design matrices for Models 1–5 (same as before)
X1 <- model.matrix(~ age_v0 + gender + dialysisvintage_v0 + diabetes_v0 + cad_v0, data = finegray_data)[, -1]
X2 <- model.matrix(~ age_v0 + gender + dialysisvintage_v0 + diabetes_v0 + cad_v0 +
                     crp_v0 + alp_v0 + albumin_v0 + phosphate_v0 + ipth_immundiagnostik_v0, data = finegray_data)[, -1]
X3 <- model.matrix(~ age_v0 + gender + dialysisvintage_v0 + diabetes_v0 + cad_v0 +
                     crp_v0 + alp_v0 + albumin_v0 + phosphate_v0 + ipth_immundiagnostik_v0 +
                     noxpthdil_v0 + beta_crosslaps_v0 + opg_v0 + lmanox_v0 + perox_v0 + srankl_v0, data = finegray_data)[, -1]

X4 <- model.matrix(~ age_v0 + gender + dialysisvintage_v0 + diabetes_v0 + cad_v0 +
                     crp_v0 + alp_v0 + albumin_v0 + phosphate_v0 + ipth_immundiagnostik_v0 +
                     cluster3_vs_rest, data = finegray_data)[, -1]
X5 <- model.matrix(~ age_v0 + gender + dialysisvintage_v0 + diabetes_v0 + cad_v0 +
                     crp_v0 + alp_v0 + albumin_v0 + phosphate_v0 + ipth_immundiagnostik_v0 +
                     noxpthdil_v0 + beta_crosslaps_v0 + opg_v0 + lmanox_v0 + perox_v0 + srankl_v0 +
                     cluster3_vs_rest, data = finegray_data)[, -1]

# 5. Fit models
fg_model1 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X1)
fg_model2 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X2)
fg_model3 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X3)
fg_model4 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X4)
fg_model5 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X5)

# and write summaries
summary(fg_model1)
summary(fg_model2)
summary(fg_model3)
summary(fg_model4)
summary(fg_model5)
```
