---
title: "cluster_outcomes"
output: html_document
author: Simon Aberger
date: "2025-10-23"
---


## Comparison of outcomes and covariates between Clusters

# Import and list relevant libraries
```{r echo=FALSE}
library(survival)
library(survminer)
library(rstatix)
library(tidyverse)
library(cmprsk)
library(MASS)
library(viridis)
```

# Load results
```{r}
load("cluster_results.RData")
```

## Step 1: Kaplan-Meier Survival Analysis (cesnored for KT) between PAM clusters
```{r}
# 1. Get events
oneyear_analysisdata <- oneyear_analysisdata %>%
  mutate(
    time_to_event = pmin(tto_death1, tto_kt1, 365, na.rm = TRUE),
    event = if_else(death_v1 == 1, 1, 0)
  )

# 2. Relabel cluster levels (for legend clarity)
oneyear_analysisdata$cluster_pamGOWER <- factor(
  oneyear_analysisdata$cluster_pamGOWER,
  labels = paste("Cluster", 1:length(unique(oneyear_analysisdata$cluster_pamGOWER)))
)

# and create survival object
surv_objPAM <- Surv(
  time = oneyear_analysisdata$time_to_event,
  event = oneyear_analysisdata$event
)

# 4. Fit KM model
km_fit_pam <- survfit(surv_objPAM ~ cluster_pamGOWER, data = oneyear_analysisdata)

# 5. Plot KM curve
km_plot_pam <- ggsurvplot(
  fit = km_fit_pam,
  data = oneyear_analysisdata,
  palette = c("#1b9e77", "#d95f02", "red", "#7570b3"),
  legend.title = "Cluster",
  legend.labs = levels(oneyear_analysisdata$cluster_pamGOWER),
  risk.table = TRUE,
  xlim = c(0, 365),                         # Limit to 365 days
  break.time.by = 90,                       # Tick marks at 0, 90, 180, 270, 360
  censor = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  xlab = "Time (days)",
  ylab = "Survival Probability",
  title = "Cumulative Survival by Cluster (Censored for KT)",
  risk.table.height = 0.35,
  risk.table.fontsize = 3.5,
  ggtheme = theme_minimal(base_size = 14) +
    theme(
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.text = element_text(color = "black"),
      axis.title = element_text(color = "black")
    )
)

km_plot_pam$plot <- km_plot_pam$plot +
  scale_x_continuous(breaks = c(0, 90, 180, 270, 365)) # Adjust x-axis ticks

print(km_plot_pam) # Show plot
```


## Step 2: CIF for secondary outcomes with competing events:
```{r}
# 1. All-cause death
oneyear_analysisdata <- oneyear_analysisdata %>%
  mutate(
    fstatus = case_when(
      death_v1 == 1 ~ 1,             # any cause of death
      transplant_v1 == 1 ~ 0,        # censored at transplant
      TRUE ~ 0
    ),
    ftime = pmin(tto_death1, tto_kt1, 365, na.rm = TRUE)
  )

cuminc_death <- cuminc(ftime   = oneyear_analysisdata$ftime, fstatus = oneyear_analysisdata$fstatus, group   = oneyear_analysisdata$cluster_pamGOWER)
print(cuminc_death)
```

# This process is repeated for `Cause-specific death`, `MACE`, `Interventions for ASCVD`, `Hospitalisation for Heart failure`, `Hospitalisation for infections` and `Fractures`



## Step 3: Incremental prediction model to test impact of cluster membership on the primary endpoint (death)
```{r}
# 1. Select model candidate variable names (by Akaike Information Criterion)
vars <- c("age_v0", "sex_v0", "dialysisvintage_v0", "bloodflow_v0", "ktv_v0",
          "dialysisdose_v0", "residualdiuresis_v0", "dialysisaccess_v0", "diabetes_v0",
          "cad_v0", "pad_v0", "heartfailure_v0", "copd_v0", "crp_v0", "alp_v0",
          "albumin_v0", "phosphate_v0", "hb_v0", "ferritin_v0", "ldl_v0")

# Start with empty model
selected <- c()
best_AIC <- Inf

AIC_crr <- function(model) -2 * model$loglik + 2 * length(model$coef)

for (v in vars) {
  # Build model adding one variable to current set
  covariates <- paste(c(selected, v), collapse = " + ")
  formula_text <- paste0("~ ", covariates)
  X <- model.matrix(as.formula(formula_text), data = finegray_data)[, -1]
  
  model <- crr(finegray_data$time_fg, finegray_data$status_fg, cov1 = X)
  new_AIC <- AIC_crr(model)
  
  cat(v, ":", new_AIC, "\n")
  
  if (new_AIC < best_AIC - 2) {
    cat("✅ Keeping", v, "(AIC improved)\n\n")
    selected <- c(selected, v)
    best_AIC <- new_AIC
  } else {
    cat("❌ Dropping", v, "(no AIC gain)\n\n")
  }
}

cat("Final selected variables:", paste(selected, collapse = ", "), "\n")
```
# A minimum improvement in AIC by 2 was elected as adequate to keep the variable in the model.

```{r}
# 2. Prepare Fine-Gray time and event variables
oneyear_analysisdata <- oneyear_analysisdata %>%
  mutate(
    status_fg = case_when(
      death_v1 == 1 ~ 1,  # Death = event
      !is.na(tto_kt1) & tto_kt1 <= 365 ~ 2,  # KT = competing risk
      TRUE ~ 0  # Censored
    ),
    time_fg = pmin(tto_death1, tto_kt1, 365, na.rm = TRUE)
  )

# and add dummy for Cluster 3 
finegray_data <- oneyear_analysisdata %>%
  mutate(
    cluster3_vs_rest = {
      cl <- as.character(cluster_pamGOWER)
      as.integer(grepl("(^3$)|3$", cl) | grepl("^Cluster\\s*3$", cl))
    }
  )

# 3.Design matrices for Models 1–5
X1 <- model.matrix(~ age_v0 + dialysisvintage_v0 + heartfailure_v0, data = finegray_data)[, -1]
X2 <- model.matrix(~ age_v0 + dialysisvintage_v0 + heartfailure_v0 +
                     albumin_v0 + phosphate_v0 , data = finegray_data)[, -1]
X3 <- model.matrix(~ age_v0 + dialysisvintage_v0 + heartfailure_v0 +
                     albumin_v0 + phosphate_v0 + 
                     crp_v0 + noxpth_v0 + beta_crosslaps_v0 + opg_v0 + lmanox_v0 + perox_v0 + 
                     srankl_v0, data = finegray_data)[, -1]

X4 <- model.matrix(~ age_v0 + dialysisvintage_v0 + heartfailure_v0 +
                     albumin_v0 + phosphate_v0 + 
                     cluster3_vs_rest, data = finegray_data)[, -1]
X5 <- model.matrix(~ age_v0 + dialysisvintage_v0 + heartfailure_v0 +
                     albumin_v0 + phosphate_v0 + 
                     crp_v0 + noxpth_v0 + beta_crosslaps_v0 + opg_v0 + lmanox_v0 + perox_v0 + 
                     srankl_v0 +
                     cluster3_vs_rest, data = finegray_data)[, -1]
# 5. Fit models
fg_model1 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X1)
fg_model2 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X2)
fg_model3 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X3)
fg_model4 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X4)
fg_model5 <- crr(ftime = finegray_data$time_fg, fstatus = finegray_data$status_fg, cov1 = X5)

# and write summaries
summary(fg_model1)
summary(fg_model2)
summary(fg_model3)
summary(fg_model4)
summary(fg_model5)
```

```{r}
# 6. Extract AIC from cmprsk::crr objects
AIC_crr <- function(model) {
  ll <- as.numeric(model$loglik)          # scalar log-likelihood
  k  <- length(model$coef)                # number of parameters
  if (is.finite(ll) && is.finite(k)) -2*ll + 2*k else NA_real_
}

aic_values <- data.frame(
  Model = c("fg_model1","fg_model2","fg_model3","fg_model4","fg_model5"),
  AIC   = c(
    AIC_crr(fg_model1),
    AIC_crr(fg_model2),
    AIC_crr(fg_model3),
    AIC_crr(fg_model4),
    AIC_crr(fg_model5)
  )
)

# Add ΔAIC and Akaike weights for interpretation
aic_values$DeltaAIC <- aic_values$AIC - min(aic_values$AIC, na.rm = TRUE)
aic_values$Weight   <- {
  w <- exp(-0.5 * aic_values$DeltaAIC)
  w / sum(w, na.rm = TRUE)
}

aic_values[order(aic_values$AIC), ]
```
